### 2021-06-10 README.Rmd
update README

---
output: github_document
always_allow_html: TRUE
---

# `peacesciencer`: Tools and Data for Quantitative Peace Science

```{r echo=FALSE, results="hide", message=FALSE}
library("badger")
library("tidyverse")
library(kableExtra)
library(peacesciencer)
library(stevemisc)
```

```{r, echo = FALSE, results='asis'}
cat(
	badge_cran_release("peacesciencer", "green"),
	badge_cran_download("peacesciencer", "grand-total", "green"),
	badge_cran_download("peacesciencer", "last-month", "green"),
	badge_cran_download("peacesciencer", "last-week", "green")
)
```

<img src="http://svmiller.com/images/peacesciencer-hexlogo.png" alt="peacesciencer  hexlogo" align="right" width="200" style="padding: 0 15px; float: right;"/>

`peacesciencer` is an R package including various functions and data sets to allow easier analyses in the field of quantitative peace science. The goal is to provide an R package that reasonably approximates what made [EUGene](https://journals.sagepub.com/doi/abs/10.1177/0738894211413055) so attractive to scholars working in the field of quantitative peace science in the early 2000s. EUGene shined because it encouraged replications of conflict models while having the user also generate data from scratch. Likewise, this R package will offer tools to approximate what EUGene did within the R environment (i.e. not requiring Windows for installation).

# Installation

You can install this on CRAN, as follows:

```r
install.packages("peacesciencer")

```

You can install the development version of this package through the `devtools` package. The development version of the package invariably has more goodies, but may or may not be at various levels of stress-testing.

```r
devtools::install_github("svmiller/peacesciencer")
```

# What's Included in `{peacesciencer}`

The package is already well developed and its functionality continues to expand. The current development version has the following functions.

```{r, echo=F, message=F, warning=F}
peacesciencer_functions <- tibble(x = as.vector(lsf.str("package:peacesciencer")))

db <- tools::Rd_db("peacesciencer")
db_names <- names(db)
db_descript <- lapply(db,function(x) {
    tags <- tools:::RdTags(x)
    if("\\description" %in% tags){
      out <- paste(unlist(x[which(tags=="\\description")]),collapse="")
    }
    else
      out <- NULL
    invisible(out)
  })

tibble(x = db_names, y = gsub("\n","",unlist(db_descript))) %>%
  mutate(x = str_sub(x, 1, nchar(x)-3)) %>%
  left_join(peacesciencer_functions, .) %>%
  mutate(func = paste0(x,"()")) -> peacesciencer_functions

peacesciencer_functions %>%
  mutate(link = paste0("[",x,"()](http://svmiller.com/peacesciencer/reference/",x,".html)")) -> peacesciencer_functions

tibble(`**Function**` = peacesciencer_functions$link,
       `**Description**` = peacesciencer_functions$y) %>%
  arrange(`**Function**`) %>%
  kable(., format="markdown")

```

The current development version also includes the following data.

```{r, echo=F, message=F, warning=F}
# library(tidyverse)
# library(kableExtra)
peacesciencer_data <- data(package = "peacesciencer")

object_name <- paste0("`",peacesciencer_data$results[, "Item"],"`")
object_link <- paste0("[",object_name,"](http://svmiller.com/peacesciencer/reference/",peacesciencer_data$results[, "Item"],".html)")

tibble(`**Object Name**` = object_link,
       `**Description**` = peacesciencer_data$results[, "Title"]) %>%
  arrange(`**Object Name**`) %>%
  kable(., format="markdown")

```

<!-- The package is very much a work in progress. Right now, it has the following functions: -->

<!-- - `add_archigos()`: adds some summary variables from Archigos about political leaders for state-year or dyad-year data. -->
<!-- - `add_atop_alliance()`: adds ATOP alliance information to dyad-year data. -->
<!-- - `add_capital_distance()`: adds capital-to-capital distance (in kilometers, "as the crow flies") to dyad-year or state-year data. -->
<!-- - `add_ccode_to_gw()`: 	Add Correlates of War state system codes to dyad-year or state-year data with Gleditsch-Ward state codes. -->
<!-- - `add_contiguity()`: adds Correlates of War direct contiguity data to dyad-year or state-year data. -->
<!-- - `add_cow_alliance()`: adds Correlates of War major alliance information to dyad-year data. -->
<!-- - `add_cow_majors()`: adds Correlates of War major power information to dyad-year or state-year data. -->
<!-- - `add_cow_mids()`: adds Correlates of War (CoW) Militarized Interstate Dispute (MID) data to dyad-year data frame -->
<!-- - `add_cow_trade()`: adds Correlates of War trade data to dyad-year or state-year data. -->
<!-- - `add_creg_fractionalization()`: add fractionalization/polarization estimates from CREG to your dyad-year or state-year data -->
<!-- - `add_democracy()`: adds estimates of democracy/levels of democracy to dyad-year or state-year data. -->
<!-- - `add_gwcode_to_cow():` adds Gleditsch-Ward state codes to dyad-year or state-year data with Correlates of War state codes. -->
<!-- - `add_gml_mids():` adds dyad-year information about ongoing MIDs and MID onsets from the Gibler-Miller-Little data. -->
<!-- - `add_igos()`: adds Correlates of War International Governmental Organizations (IGOs) data to dyad-year or state-year data. -->
<!-- - `add_mids():` adds dyad-year information about ongoing MIDs and MID onsets from the Gibler-Miller-Little data. (DEPRECATED) -->
<!-- - `add_minimum_distance()`: adds minimum distance (in kilometers) to dyad-year or state-year data. -->
<!-- - `add_nmc()`: adds estimates of national material capabilities (from Correlates of War) to dyad-year or state-year data. -->
<!-- - `add_peace_years()`: Add Peace Years to Your Conflict Data -->
<!-- - `add_sdp_gdp()`: adds estimates of (gross and surplus) domestic product and population size to dyad-year or state-year data -->
<!-- - `add_strategic_rivalries()`: Add Thompson and Dreyer's (2012) strategic rivalry data to dyad-year data frame -->
<!-- - `add_ucdp_acd()`: Add UCDP Armed Conflict Data to state-year data frame -->
<!-- - `add_ucdp_onsets()`: adds UCDP onsets to state-year data -->
<!-- - `create_dyadyears()`: converts Correlates of War or Gleditsch-Ward state system membership data into dyad-year format -->
<!-- - `create_statedays()`: converts Correlates of War or Gleditsch-Ward state membership data into state-day format.. -->
<!-- - `create_stateyears()`: converts Correlates of War or Gleditsch-Ward state membership data into state-year format. -->
<!-- - `filter_prd()`: filters dyad-year data frame to just "politically relevant" dyads. -->

<!-- It also has the following data sets: -->

<!-- - `archigos`: an abbreviated version of the Archigos data, used internally -->
<!-- - `atop_alliance:` directed dyad-year alliance data from ATOP -->
<!-- - `capitals`: a list of capitals and capital transitions for Correlates of War state system members -->
<!-- - `ccode_democracy`: Correlates of War state-year data with three different estimates of democracy (Varieties of Democracy, Polity, Xavier Marquez/Pemstein et al.'s "(Quick) Unified Democracy Scores"). -->
<!-- - `cow_alliance`: directed dyad-year alliance formation data from the Correlates of War -->
<!-- - `cow_contdir`: Correlates of War Direct Contiguity Data (v. 3.2) -->
<!-- - `cow_ddy`: a full directed dyad-year data frame of Correlates of War state system members -->
<!-- - `cow_gw_years`: a yearly data frame including information about Correlates of War and Gleditsch-Ward states. -->
<!-- - `cow_igo_ndy`: non-directed dyad-year data for Correlates of War intergovernmental organizations data. -->
<!-- - `cow_igo_sy`: state-year data for Correlates of War intergovernmental organizations data. -->
<!-- - `cow_majors`: Correlates of War major powers data (version: 2016) -->
<!-- - `cow_mid_ddydisps`:  Directed Dyadic Dispute-Year Data with No Duplicate Dyad-Years (CoW-MID, v. 5.0) -->
<!-- - `cow_mid_dirdisps`: Correlates of War Militarized Interstate Dispute data (directed dyadic dispute-year, v. 5.0)  -->
<!-- - `cow_mid_disps`: Correlates of War Militarized Interstate Dispute data (abbreviated dispute-level, v. 5.0)  -->
<!-- - `cow_mindist`: the minimum distance between Correlates of War states, in kilometers, in non-directed dyad-year format (1946-2015) -->
<!-- - `cow_nmc`: Correlates of War National Material Capabilities data (version 5.0) -->
<!-- - `cow_sdp_gdp`: (Surplus and Gross) Domestic Product for Correlates of War States -->
<!-- - `cow_states`: Correlates of War state system membership data (version: 2016) -->
<!-- - `cow_trade_sy`: Correlates of War state-year trade data (version 4.0) -->
<!-- - `creg`: Composition of Religious and Ethnic Groups (CREG) Fractionalization/Polarization Estimates -->
<!-- - `gml_dirdisp`: directed dispute-year data from version 2.1.1 of the Gibler-Miller-Little inter-state dispute data. -->
<!-- - `gml_mid_ddydisps`:  Directed Dyadic Dispute-Year Data with No Duplicate Dyad-Years (Gibler-Miller-Little, v. 2.1.1) -->
<!-- - `gw_ddy`: a full directed dyad-year data frame of Gleditsch-Ward state system members -->
<!-- - `gw_mindist`: the minimum distance between  Gleditsch-Ward states, in kilometers, in non-directed dyad-year format (1946-2015) -->
<!-- - `gw_sdp_gdp`: (Surplus and Gross) Domestic Product for Correlates of War States -->
<!-- - `gw_states`: Gleditsch-Ward independent state system data (version: 2017) -->
<!-- - `gwcode_democracy`: Gleditsch-Ward state-year data with three different estimates of democracy (Varieties of Democracy, Polity, Xavier Marquez/Pemstein et al.'s "(Quick) Unified Democracy Scores"). -->
<!-- - `hief`: Historical Index of Ethnic Fractionalization data -->
<!-- - `maoz_powers`: Zeev Maoz' global/regional power data -->
<!-- - `rugged`: Rugged/Mountainous Terrain Data -->
<!-- - `td_rivalries`: Thompson and Dreyer's (2012) Strategic Rivalries, 1494-2010 -->
<!-- - `ucdp_acd`: a (not quite) dyad-year and (not quite) state-year data set on armed conflict episodes -->
<!-- - `ucdp_onsets`: a state-year data set of UCDP armed conflict onsets/episodes. -->

# How to Use `{peacesciencer}`

[`{peacesciencer}` has a user's guide](https://github.com/svmiller/peacesciencer/blob/master/user-guide/doc/ms.pdf) that is worth reading. The workflow is going to look something like this. This is a "tidy"-friendly approach to a data-generating process in quantitative peace science.

First, start with one of two processes to create either dyad-year or state-year data. The dyad-year data are created with the `create_dyadyears()` function. It has a few optional parameters with hidden defaults. The user can specify what kind of state system (`system`) data they want to use---either Correlates of War (`"cow"`) or Gleditsch-Ward (`"gw"`), whether they want to extend the data to the most recently concluded calendar year (`mry`) (i.e. Correlates of War state system membership data are current as of Dec. 31, 2016 and the script can extend that to the end of the most recently concluded calendar year), and whether the user wants directed or non-directed dyad-year data (`directed`).

The `create_stateyears()` works much the same way, though "directed" and "non-directed" make no sense in the state-year context. Both functions default to Correlates of War state system membership data to the most recently concluded calendar year.

Thereafter, the user can specify what additional variables they want added to these dyad-year or state-year data. Do note: the additional functions lean primarily on Correlates of War state code identifiers. Indeed, the bulk of the quantitative peace science data ecosystem is built around the Correlates of War project. The variables the user wants are added in [a "pipe"](https://r4ds.had.co.nz/pipes.html) in a process like this. Do note that the user may want to break up the data-generating process into a few manageable "chunks" (e.g. first generating dyad-year data and saving to an object, adding to it piece by piece).

All told, the process will look something like this. Assume you want to create some data for something analogous to a "dangerous dyads" design for all non-directed dyad-years. Here's how you'd do it in `{peacesciencer}`, which is going to be lifted from the source R scripts for the user's guide. The first part of this code chunk will lean on core `{peacesciencer}` functionality whereas the other stuff is some post-processing and, as a bonus, some modeling.

```{r, cache=F, message=F, eval=T}

#library(tidyverse)
#library(peacesciencer)
# library(stevemisc)
library(tictoc)

tic()
create_dyadyears(directed = FALSE, mry = FALSE) %>%
  filter_prd() %>%
  add_gml_mids(keep = NULL) %>%
  add_peace_years() %>%
  add_nmc() %>%
  add_democracy() %>%
  add_cow_alliance() %>%
  add_sdp_gdp() -> Data

Data %>%
  mutate(landcontig = ifelse(conttype == 1, 1, 0)) %>%
  mutate(cowmajdyad = ifelse(cowmaj1 == 1 | cowmaj2 == 1, 1, 0)) %>%
  # Create estimate of militarization as milper/tpop
  # Then make a weak-link
  mutate(milit1 = milper1/tpop1,
         milit2 = milper2/tpop2,
         minmilit = ifelse(milit1 > milit2,
                           milit2, milit1)) %>%
  # create CINC proportion (lower over higher)
  mutate(cincprop = ifelse(cinc1 > cinc2,
                           cinc2/cinc1, cinc1/cinc2)) %>%
  # create weak-link specification using Quick UDS data
  mutate(mindemest = ifelse(xm_qudsest1 > xm_qudsest2,
                            xm_qudsest2, xm_qudsest1)) %>%
  # Create "weak-link" measure of jointly advanced economies
  mutate(minwbgdppc = ifelse(wbgdppc2011est1 > wbgdppc2011est2,
                             wbgdppc2011est2, wbgdppc2011est1)) -> Data

# r2sd() is in {stevemisc}, a {peacesciencer} dependency.
Data %>%
  mutate_at(vars("cincprop", "mindemest", "minwbgdppc", "minmilit"),
            ~r2sd(.)) -> Data

summary(modDD <- glm(gmlmidonset ~ landcontig + cincprop + cowmajdyad + cow_defense +
               mindemest + minwbgdppc + minmilit +
               gmlmidspell + I(gmlmidspell^2) + I(gmlmidspell^3), data= Data,
             family=binomial(link="logit")))
toc()

```

Here is how you might do a standard civil conflict analysis using Gleditsch-Ward states and UCDP conflict data.

```{r, cache=F, message=F, eval=T}

tic()
create_stateyears(system = 'gw') %>%
  filter(year %in% c(1946:2019)) %>%
  add_ucdp_acd(type=c("intrastate"), only_wars = FALSE) %>%
  add_peace_years() %>%
  add_democracy() %>%
  add_creg_fractionalization() %>%
  add_sdp_gdp() %>%
  add_rugged_terrain() -> Data

create_stateyears(system = 'gw') %>%
  filter(year %in% c(1946:2019)) %>%
  add_ucdp_acd(type=c("intrastate"), only_wars = TRUE) %>%
  add_peace_years() %>%
  rename_at(vars(ucdpongoing:ucdpspell), ~paste0("war_", .)) %>%
  left_join(Data, .) -> Data

Data %>%
  arrange(gwcode, year) %>%
  group_by(gwcode) %>%
  mutate_at(vars("xm_qudsest", "wbgdppc2011est",
                 "wbpopest"), list(l1 = ~lag(., 1))) %>%
  rename_at(vars(contains("_l1")),
            ~paste("l1", gsub("_l1", "", .), sep = "_") ) -> Data

modCW <- list()
summary(modCW$"All UCDP Conflicts" <- glm(ucdponset ~ l1_wbgdppc2011est + l1_wbpopest  +
                    l1_xm_qudsest + I(l1_xm_qudsest^2) +
                    newlmtnest + ethfrac + relfrac +
                    ucdpspell + I(ucdpspell^2) + I(ucdpspell^3), data=subset(Data),
                  family = binomial(link="logit")))

summary(modCW$"Wars Only"  <- glm(war_ucdponset ~ l1_wbgdppc2011est + l1_wbpopest  +
                    l1_xm_qudsest + I(l1_xm_qudsest^2) +
                    newlmtnest + ethfrac + relfrac +
                    war_ucdpspell + I(war_ucdpspell^2) + I(war_ucdpspell^3), data=subset(Data),
                  family = binomial(link="logit")))

toc()
```

# Issues/Requests

`{peacesciencer}` is already more than capable to meet a wide variety of needs in the peace science community. Users are free to raise an issue on the project's Github if some feature is not performing as they think it should or if there are additions they would like to see.