### 2021-11-02 data-raw/informal-tests/informal-tests.Rmd
update. I think the leader contributions are sufficiently tested.

---
title: "More Tests"
author: "steve"
date: "`r Sys.Date() `"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Check if Stuff is Already Downloaded

```{r}
library(tidyverse)
library(peacesciencer)
library(testthat)
download_extdata()
```

## Create Some Data

```{r}
cow_leaderyears <- create_leaderyears(standardize_cow = TRUE)
gw_leaderyears <- create_leaderyears(standardize_cow = FALSE)

cow_leaderdyadyears <- create_leaderdyadyears(system = 'cow')
gw_leaderdyadyears <- create_leaderdyadyears(system = 'gw')

cow_leadermids <- cow_leaderyears %>% add_gml_mids() %>% filter(year %in% c(1875:2010))
cow_leaderdyadmids <- cow_leaderdyadyears %>% add_gml_mids() %>% filter(year %in% c(1875:2010))
```

## Run Some Tests

These are tests that add data based on what's stored remotely. I don't trust CRAN to do these tests well---and already so much is tested---so I'll communicate the results of those tests here. The leader-year tests are here too. While these could be done as part of the formal testing process, they would also take forever and the time from `devtools::check()` to `devtools::release()` is already like half an hour or more. In due time, these leader-year tests could---and probably should---go into the formal testing procedure.

```{r}
test_that("Dyad-year additions do not create duplicates", {
  expect_equal(nrow(cow_ddy), nrow(cow_ddy %>% add_cow_trade()))
})
```

```{r}
test_that("Leader-year additions do not create duplicates", {
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_capital_distance()))
  expect_equal(nrow(gw_leaderyears), nrow(gw_leaderyears %>% add_capital_distance()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_contiguity()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_cow_majors()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_cow_trade()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_creg_fractionalization()))
  expect_equal(nrow(gw_leaderyears), nrow(gw_leaderyears %>% add_creg_fractionalization()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_democracy()))
  expect_equal(nrow(gw_leaderyears), nrow(gw_leaderyears %>% add_democracy()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_gml_mids()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_minimum_distance()))
  expect_equal(nrow(gw_leaderyears), nrow(gw_leaderyears %>% add_minimum_distance()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_nmc()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_rugged_terrain()))
  expect_equal(nrow(gw_leaderyears), nrow(gw_leaderyears %>% add_rugged_terrain()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_sdp_gdp()))
  expect_equal(nrow(gw_leaderyears), nrow(gw_leaderyears %>% add_sdp_gdp()))
  expect_equal(nrow(cow_leadermids), nrow(cow_leadermids %>% add_spells()))
  expect_equal(nrow(cow_leaderyears), nrow(cow_leaderyears %>% add_gwcode_to_cow()))
  expect_equal(nrow(gw_leaderyears), nrow(gw_leaderyears %>% add_ccode_to_gw()))
})
```

```{r}
test_that("Leader-dyad-year additions do not create duplicates", {
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_capital_distance()))
  expect_equal(nrow(gw_leaderdyadyears), nrow(gw_leaderdyadyears %>% add_capital_distance()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_contiguity()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_cow_majors()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_cow_trade()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_creg_fractionalization()))
  expect_equal(nrow(gw_leaderdyadyears), nrow(gw_leaderdyadyears %>% add_creg_fractionalization()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_democracy()))
  expect_equal(nrow(gw_leaderdyadyears), nrow(gw_leaderdyadyears %>% add_democracy()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_gml_mids()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_minimum_distance()))
  expect_equal(nrow(gw_leaderdyadyears), nrow(gw_leaderdyadyears %>% add_minimum_distance()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_nmc()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_rugged_terrain()))
  expect_equal(nrow(gw_leaderdyadyears), nrow(gw_leaderdyadyears %>% add_rugged_terrain()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_sdp_gdp()))
  expect_equal(nrow(gw_leaderdyadyears), nrow(gw_leaderdyadyears %>% add_sdp_gdp()))
  expect_equal(nrow(cow_leaderdyadmids), nrow(cow_leaderdyadmids %>% add_spells()))
  expect_equal(nrow(cow_leaderdyadyears), nrow(cow_leaderdyadyears %>% add_gwcode_to_cow()))
  expect_equal(nrow(gw_leaderdyadyears), nrow(gw_leaderdyadyears %>% add_ccode_to_gw()))
})
```