### 2021-11-05 user-guide/memo.Rmd

---
# output: 
#   stevetemplates::cover_letter:
#     latex_engine: xelatex 
geometry: margin=1in
header-includes:
  - \linespread{1.05}

author: "&nbsp;"
address: ""
email: ""
url: ""
phone: ""
fontsize: 11pt
mainfont: cochineal
monofont: Fira Code

fontawesome: TRUE
date: "`r format(Sys.time(), '%e %B %Y')`"
greetings: "Dear Editor and Reviewers,"
bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`"
biblio-style: apsr
---

```{r setup, include=F}

```

I thank you all for the opportunity to submit revisions of my manuscript "Asfsfssdfsafdsafs3fs" for possible inclusion in *Conflict Management and Peace Science* (*CMPS*). I thank the two reviewers and the editorial team for the helpful comments as well.

# Comments to the Editorial Team

The editorial team noted

Enclosed are the revised manuscript and appendix. I thank the editorial team and the reviewers for their helpful comments and for further consideration of my revisions.

Best regards,

```{r eval=F, echo=FALSE, out.width='30%'}
# knitr::include_graphics('john-hancock.png')
# ^ put your signature here as a PNG or something.
```

The Author

\newpage

# Reviewer #1's Comments

Reviewer #1 (R1) had the following comments on the original version of the manuscript.

\newpage

# Reviewer #2's Comments

I thank Reviewer #2 (R2) for

\newpage